This script is the Tidymodels workflow for experimenting with multiple model types trained to predict rust tide occurrences on the Narragansett Bay Long Term Time Series. 

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
source(get_path(SCRIPT_DIR, "NarragansettLong", "NarLongDF.R"))
```

Reading in required packages
```{r}
library(tidyverse)
library(tidymodels)
library(themis)
library(vip)
library(unpivotr)
```

Creating the dataset for modeling with lags and factorizing the count values
```{r}
set.seed(15)
data <- count_env |>
  mutate(woy = week(Date)) |>
  mutate(flownolag = flow_mean) |>
  mutate(count = as.factor(
    ifelse(`count` >= 5000, 'high', "low" )))|>
  mutate_at(c("count", "id", "woy"),lead, n = 1) |>
  mutate_at(c("Surface Temp"), lag, n = 1) |>
  mutate_at(c("Surface Salinity"), lag, n = 1) |>
  mutate_at(c("TotPAR"), lag, n = 0) |>
  mutate_at(c("flow_mean"), lag, n = 2) |>
  mutate_at(c("WSpd"), lag, n = 3) |>
  mutate_at(c("Wdir"), lag, n = 3) |>
  mutate_at(c("BP"), lag, n = 0) 
```

Creating training and testing sets with desired covariates
```{r}
#splitting training and testing data sets
set.seed(15)
split <- initial_time_split(data, prop = 9/10, lag = 0)
train_raw <- training(split)
test_raw <- testing(split)

test <- test_raw |>
  dplyr::select( "count", "Surface Temp", "TotPAR", "flow_mean", "Wdir", "Surface Salinity")|>
  tidyr::drop_na()

train <- train_raw |>
  dplyr::select( "count" ,"Surface Temp", "TotPAR", "flow_mean", "Wdir", "Surface Salinity")|>
  tidyr::drop_na()
```

Defining the models to test
```{r}
mlnn <- mlp(mode = "classification", engine = "brulee", hidden_units = 4, epochs = 100, dropout = .05, activation = "relu", learn_rate = .002) 
rf <- rand_forest(mode = "classification", engine = "ranger", trees = 2000, mtry = 3) 
bag <- bag_tree(
  mode = "classification",
  cost_complexity = 0,
  tree_depth = NULL,
  min_n = 2,
  class_cost = NULL,
  engine = "rpart"
)
```

Defining evaluation metrics and folds for cross validations 
```{r}
resamples = vfold_cv(train, v = 5, strata = count)

model_metrics <- yardstick::metric_set( yardstick::accuracy,  yardstick::sensitivity,  yardstick::specificity, yardstick::precision,  yardstick::recall,  yardstick::f_meas)
```

Creating processing recipe 
```{r}
set.seed(15)
rust_recipe <- train |> recipe(count ~ ., data = train)  |>
  step_log(TotPAR, signed = TRUE)|>
  step_log(flow_mean)|>
  step_BoxCox() |>
  step_normalize(all_numeric_predictors()) |>
  step_dummy(all_nominal_predictors()) |>
  step_upsample(count, over_ratio = .35) 
  
```

Creating a workflow
```{r}
bal_wf <-
  workflow() |>
  add_recipe(rust_recipe) |>
  add_model(mlnn)
```


Tuning model
```{r eval=FALSE, include=FALSE}

  mlp_spec <-
    mlp(hidden_units = tune(), epochs = tune()) |>
    set_mode("classification")


ames_mlp_itr <-
    mlp_spec %>%
    tune_bayes(
      rust_recipe,
      resamples = resamples,
      initial = 20,
      iter = 4,
      control = control_bayes(save_pred = TRUE)
    )


```

Validation set testing 
```{r}
set.seed(15)
bal_rs <- 
  bal_wf |>
  fit_resamples(
    resamples = resamples,
    control = control_resamples(save_pred = TRUE)) 

rs_confusionmatrix <- conf_mat_resampled(bal_rs)

rs_metrics <-   fit_resamples(bal_wf,
    resamples = resamples, metrics = model_metrics) |>
  collect_metrics()
```

Fitting,predicting, and getting model metrics
```{r}
set.seed(15)
fit <- fit(bal_wf, data = train)
pred <- augment(fit, test)
dat_met <- metrics(pred, truth = `count`, estimate=.pred_class)
```

Creating a confusion matrix for test set results 
```{r}
confusion_matrix_test <- conf_mat(pred, truth = count, estimate = .pred_class)
plt <- as.data.frame(confusion_matrix_test$table)
confusionmat <- ggplot(plt, aes(Prediction,Truth, fill= Freq)) +
  geom_tile() + geom_text(aes(label=Freq)) +
  scale_fill_gradient(low="white", high="#009194") +
  labs(x = "Prediction",y = "Truth") +
  scale_x_discrete(labels=c("High","Low")) +
  scale_y_discrete(labels=c("High","Low"))
```

Graphing variable importance
```{r eval=FALSE, include=FALSE}
varimp <- fit |>
  extract_fit_parsnip() |>
  vip(type = "garson")
```

Making plots of model results and model structure
```{r}
set.seed(15)
data_conf <- count_env |>
  mutate(flownolag = flow_mean) |>
  mutate_at(c("count", "id"),lead, n = 1) |>
  mutate_at(c("Surface Temp"), lag, n = 1) |>
  mutate_at(c("Surface Salinity"), lag, n = 1) |>
  mutate_at(c("TotPAR"), lag, n = 0) |>
  mutate_at(c("flow_mean"), lag, n = 2) |>
  mutate_at(c("WSpd"), lag, n = 3) |>
  mutate_at(c("Wdir"), lag, n = 3) |>
  mutate_at(c("BP"), lag, n = 0)

split <- initial_time_split(data_conf, prop = 9/10, lag = 0)
confcounts <- testing(split)


countcomp <- confcounts |>
  dplyr::select( "count", "Surface Temp", "TotPAR", "flow_mean","Wdir", "Date", "Surface Salinity") |>
  drop_na() 


countcomp <- countcomp |>
  cbind(pred$.pred_high) 
  for(i in 1:length(countcomp$count)){
    if(countcomp$count[i]>0){
      countcomp$count[i] = log10(countcomp$count[i])
    }
  }


ggplot(data = countcomp) +
  geom_jitter(aes(x = pred$.pred_high, y = count), width = .01, color = "blue") +
  geom_hline(yintercept=log10(5000), linetype='dashed') +
  geom_vline(xintercept = .5, linetype = 'dashed') +
  xlab("Predicted Probability of \"High\" Class") +
  ylab("Log True Abundance (log(Cells/L))") +
  ggtitle("Confusion Matrix")+ 
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data = countcomp, aes(x = Date, y = pred$.pred_high)) +
  geom_point()
#response curves
lmfit = lm(log(countcomp$`Surface Temp`) ~ countcomp$`pred$.pred_high`,data=countcomp)
A = coef(lmfit)[1]
B = coef(lmfit)[2]

sstres <- ggplot(data = countcomp, aes(x = `Surface Temp`, y = `pred$.pred_high`)) +
  geom_point() +
  ylab("High Prediction Percentage")

library(devtools)
source_url('https://gist.githubusercontent.com/fawda123/7471137/raw/466c1474d0a505ff044412703516c34f1a4684a5/nnet_plot_update.r')

plot.nnet(extract_model(fit), y.lab = "Binary Output", x.lab = c("Surface Temperature", "Total PAR", "Mean River Flow", "Surface Salinity", "Wind Direction"), bias = FALSE, cex.val = 1.25, circle.cex = 6, line.stag = .0001)
```

